---
title: "Seider"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Load Required Packages
```{r Loading Required Packages}
library(rpart)
library(rpart.plot)
library(recipes)
library(caret)
library(ggplot2)
library(tidyverse)
library(ROCR)
library(dummies)
library(fpc)
library(clustMixType)
library(xlsx)

```
#Load Data Set
```{r Load Data Set}

Crash <- read.csv("C:/Users/Ab/Documents/MSBIA/Applied Data Mining/HW 3/Data/crash_clean.csv")
dim(Crash)
str(Crash)
summary(Crash)
```
#Create Region Variable
```{r Create Region Variable}
#Create Region V
summary(Crash$State)
Crash$Region<-ifelse(Crash$State=="Illinois"|Crash$State=="Indiana"|
                       Crash$State=="Michigan"|Crash$State=="Ohio"|
                       Crash$State=="Wisconsin"|Crash$State=="Iowa"|
                       Crash$State=="Kansas"|Crash$State=="Minnesota"|
                       Crash$State=="Missouri"|Crash$State=="Nebraska"|
                       Crash$State=="North Dakota"|Crash$State=="South Dakota","Midwest",
                     ifelse(Crash$State=="Connecticut"|Crash$State=="Maine"|
                              Crash$State=="Massachusetts"|Crash$State=="New Hampshire"|
                              Crash$State=="Rhode Island"|Crash$State=="Vermont"|
                              Crash$State=="New Jersey"|Crash$State=="New York"|
                              Crash$State=="Pennsylvania", "Northeast",
                            ifelse(Crash$State == "Delaware"| Crash$State =="Florida" | Crash$State == "Georgia" |
                                     Crash$State == "Maryland" | Crash$State == "North Carolina"
                                   | Crash$State == "South Carolina" | Crash$State == "Virginia"
                                   | Crash$State == "District of Columbia" | Crash$State == "West Virginia"
                                   | Crash$State == "Alabama" | Crash$State == "Kentucky"
                                   | Crash$State == "Mississippi" | Crash$State == "Tennessee"
                                   | Crash$State == "Arkansas" | Crash$State == "Louisiana" 
                                   | Crash$State == "Oklahoma" | Crash$State == "Texas", "South",
                                   ifelse(Crash$State == "Arizona" | Crash$State == "Colorado"
                                          | Crash$State == "Idaho" | Crash$State == "Montana"
                                          | Crash$State == "Nevada" | Crash$State == "New Mexico" 
                                          | Crash$State == "Utah" | Crash$State == "Wyoming"
                                          | Crash$State == "Alaska" | Crash$State == "California"
                                          | Crash$State == "Hawaii" | Crash$State == "Oregon"
                                          | Crash$State == "Washington","West","NA"))))

table(Crash$Region)
```
#Find Missing Values Percentage
```{r Find Missing Values Percentage}
pMissing <- function(x){sum(is.na(x))/length(x)*100} 
apply(Crash,2,pMissing)
```
#Histogram of the Age
```{r}
hist(as.numeric(Crash$Age))
```
Filter out Unknown Ages as there is not logical way to infer
```{r}
Crash <- filter (Crash, Crash$Age != "Unknown")
apply(Crash,2,pMissing)
```
#More Data Cleanup
```{r}
#Remove all of the original variable before they were cleaned in excel
names(Crash)
Crash <- Crash[,-c(5,7,9:11)]
names(Crash)
Crash <- Crash[,-c(12)]
#Rename the cleaned variable
names(Crash)
Crash$Race <- Crash$Race.Cleaned
Crash$Gender <- Crash$Gender.Cleaned
Crash$Drug.Involvement <- Crash$Drug.Involvment.Clean
Crash$Age <- Crash$Age.Cleaned
names(Crash)
#Remove the cleaned variable since the are now duplicates
Crash <- Crash[,-c(12:15)]
names(Crash)
summary(Crash)
apply(Crash,2,pMissing)
names(Crash)
#Remove date, year, month, day, week, and state variables
Crash <- Crash[, -c(1,3,8:11)]
apply(Crash,2,pMissing)
names(Crash)
```

#Z score the number variables
```{r}
#Crash$Fatalities.in.crash <- as.data.frame(scale(Crash$Fatalities.in.crash))
#Crash$Age <- as.numeric(Crash$Age)
#Crash$Age <- as.data.frame(scale(Crash$Age))
#summary(Crash$Age)
```

#create dummy variables for all categorical variables
```{r}
Crash.D <- dummy.data.frame(Crash, names = 'Atmospheric.Condition', sep = ":")
Crash.D <- dummy.data.frame(Crash.D, names = 'Roadway.Category', sep = ":")
Crash.D <- dummy.data.frame(Crash.D, names = 'Person.Type', sep = ":")
Crash.D <- dummy.data.frame(Crash.D, names = 'Injury.Severity', sep = ":")
Crash.D <- dummy.data.frame(Crash.D, names = 'Region', sep = ":")
Crash.D <- dummy.data.frame(Crash.D, names = 'Race', sep = ":")
Crash.D <- dummy.data.frame(Crash.D, names = 'Gender', sep = ":")
Crash.D <- dummy.data.frame(Crash.D, names = 'Drug.Involvement', sep = ":")
names(Crash.D)
head(Crash.D)
```

#Min-Max Normalize Numeric Feilds
```{r minMix}
MinMaxNormalize <- function(x){return((x-min(x))/(max(x)-min(x)))}
CrashNum <- Crash.D[,c(13:14)]
CrashNum$Age <- as.integer(CrashNum$Age)
CrashNormal <- as.data.frame(lapply(CrashNum, MinMaxNormalize))
summary(CrashNormal)
```
#Put Data Set Back Together
```{R cbind}
Crash.D = lapply(Crash.D, factor)
CrashModel <- cbind(Crash.D, CrashNormal)
names(CrashModel)
CrashModel <- CrashModel[,-c(13:14)]
names(CrashModel)
lapply(CrashModel, class)
```

# Elbow Method for finding the optimal number of clusters
# Compute and plot wss for k = 2 to k = 15.
```{r elbow}
k.max <- 15

wss <- sapply(1:k.max, 
              function(k){kproto(CrashModel, k)$tot.withinss})
wss
plot(1:k.max, wss,
     type="b", pch = 19, frame = FALSE, 
     xlab="Number of clusters K",
     ylab="Total within-clusters sum of squares")
```
#Creat Clusters With KProto
```{r}
set.seed(123)
```
#3 Clusters
```{r 3clusters}
CrashKProto3 <- kproto(CrashModel, k= 3)
summary(CrashKProto3)
CrashKProto3$size
ProfileKProto3 <- data.frame(t(CrashKProto3$centers))
ProfileKProto3
clprofiles(CrashKProto3, CrashModel)


```
#8 Clusters
```{r 8clusters}
CrashKProto8 <- kproto(CrashModel, k= 8)
summary(CrashKProto8)
CrashKProto8$size
ProfileKProto8 <- data.frame(t(CrashKProto8$centers))
ProfileKProto8
clprofiles(CrashKProto8, CrashModel)


```

#Cluster Cohesion
#Calculate withinss

```{r Cohesion}
KProto3C <- CrashKProto3$withinss/CrashKProto3$tot.withinss
KProto8C <- CrashKProto8$withinss/CrashKProto8$tot.withinss

KProtoCohesion <- c(KProto3C,KProto8C)
print(KProtoCohesion)
```

#Add the clusters back to the original data set
```{r AddBack}
Crash$kProto3 <- CrashKProto3$cluster
Crash$kProto8 <- CrashKProto8$cluster
```

#Export the data
```{r Export}
write.xlsx(Crash, "C:/Users/Ab/Documents/MSBIA/Applied Data Mining/HW 3/Data/CrashCluster.xlsx")
```